---
title: "ARE 213 PS 3"
author: "S. Sung, H. Husain, T. Woolley, A. Watt"
email: aaron@acwatt.net
date: "2021-12-8"
output:
  pdf_document:
    toc: true
    toc_depth: 2
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{amsmath}
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE_213_PS2b.pdf')) })
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = F)
```
<!--
R version 3.6.3 (2020-02-29)
Purpose of script: run code necessary for ps2b for ARE 213

Notes: Need to open ps2b.Rmd from the ps2b folder (without Rstudio being started first)
if Rstudio is already started, the working directory will not be set
to the ps2b/ folder

\usepackage{dcolumn}: dcolumn is needed in latex compilation for multicolumn tables
-->

```{r Settings}
# stargazer table type (html, latex, or text)
# Change to latex when outputting to PDF, html when outputting to html
table_type = "latex"
```


```{r packages, include=F}
library(tidyverse)
library(haven)
library(stargazer)
library(ggplot2)
library(tinytex)
library(lfe) # For felm() fixed effects model
library(lmtest)
library(sandwich)
# library(Synth)
# library(plm)
# library(gridExtra)
# library(grid)
# library(gtable)
# library(fastDummies)
# library(EnvStats)
```


```{r Call in all Datasets, eval=T}
allsite = read_dta('allsites.dta')
allcovariates = read_dta('allcovariates.dta')
twomiledata = read_dta('2miledata.dta')
sitecovariates = read_dta('sitecovariates.dta')
```


```{r Data Exploration - allsite.dta, eval=T}
# In "allsite.dta", we observe that there are duplicate observations. Dropped
sum(duplicated(allsite$fips)) # There are cases of duplicate observations with same FIPs.
sum(duplicated(allsite)) # This confirms that their entire row values are 
                         # duplicates and can be dropped.
# Duplicates Dropped
allsite <- allsite[!duplicated(allsite),]
```


\newpage
<!--=========================================================================-->
# Problem 1
<!--=========================================================================-->
**This question asks you to run OLS regressions that look at whether there is an association between 2000 housing values and whether a census tract contained a hazardous waste site that was placed on the NPL by 2000.**

## Part (a)
**Use the file allsites.dta. This file contains only own tract housing variables (i.e. no 2 mile averages). Use "robust" standard errors for all regressions. First regress 2000 housing prices on whether the census tract had an NPL site in 2000. Include 1980 housing values as a control. Next add housing characteristics as controls. Run a third regression adding economic and demographic variables as controls. Finally run a 4th regression that also includes state fixed effects. Briefly interpret the regressions. Under what conditions will the coeffcients on NPL 2000 status be unbiased?**

```{r Q1a}
# Regress 2000 housing prices on 1(NPL site in 2000) and 1980 housing values.
reg_1a1 <- lm(lnmdvalhs0 ~ npl2000 + lnmeanhs8, data = allsite)
reg_1a1_r <- coeftest(reg_1a1, vcov = vcovHC(reg_1a1, type = "HC1"))

# Regress: Add housing characteristics as control.
HousingChar <- c("smhse8", "tothsun8", "ownocc8", "firestoveheat80", "noaircond80", 
                 "nofullkitchen80", "zerofullbath80", 
                "bedrms1_80occ", "bedrms2_80occ", "bedrms3_80occ", "bedrms4_80occ", "bedrms5_80occ",
                "blt2_5yrs80occ", "blt6_10yrs80occ", "blt10_20yrs80occ", "blt20_30yrs80occ", "blt30_40yrs80occ", "blt40_yrs80occ",
                "detach80occ", "attach80occ", "mobile80occ", "occupied80")
Formula1 <- formula(paste("lnmdvalhs0 ~ npl2000 + lnmeanhs8 + ", paste(HousingChar, collapse=" + ")))
reg_1a2 <- lm(Formula1, data = allsite)
reg_1a2_r <- coeftest(reg_1a2, vcov = vcovHC(reg_1a2, type = "HC1"))

# Regress: Add economic and demographic variables as control.
EconNDemo <- c("pop_den8", "shrblk8", "shrhsp8", "child8", "old8", "shrfor8", "ffh8", 
               "hsdrop8", "no_hs_diploma8", "ba_or_better8", "unemprt8", "povrat8", 
               "welfare8", "avhhin8")
Formula2 <- formula(paste("lnmdvalhs0 ~ npl2000 + lnmeanhs8 + ", paste(c(HousingChar, EconNDemo), collapse=" + ")))
reg_1a3 <- lm(Formula2, data = allsite)
reg_1a3_r <- coeftest(reg_1a3, vcov = vcovHC(reg_1a3, type = "HC1"))

# Include State FE
Formula3 <- formula(paste("lnmdvalhs0 ~ npl2000 + lnmeanhs8 + ", paste(c(HousingChar, EconNDemo), collapse=" + "), " | statefips"))
reg_1a4 <- felm(Formula3, data = allsite)

# Stargazer Table with All Four Regressions
stargazer(reg_1a1_r, reg_1a2_r, reg_1a3_r, reg_1a4,
          se = list(reg_1a1_r[,2], reg_1a2_r[,2], reg_1a3_r[,2], reg_1a4$rse),
          type = "text", omit.stat = "f")

# Questions
#Why multicollinear
#Do we want clustered SE
```




\newpage
## Part (b)
**Here we will compare covariates between potential treatment and comparison groups. First, use allcovariates.dta to compare co-variates (i.e. those used in the above regressions) between census tracts with and without a hazardous waste site listed on the NPL by 2000. Next, use sitecovariates.dta to compare covariates between those census tracts with a hazardous waste site that had an HRS test in 1982. Specifically, compare those with sites that scored above 28.5 to those that scored below 28.5. Finally, compare those census tracts with sites between 16.5 and 28.5 to census tracts with sites between 28.5 and 40.5. What conclusions do you draw from these 3 comparisons?**






\newpage
<!--=========================================================================-->
# Problem 2
<!--=========================================================================-->
**This question examines the possibility of using a Regression Discontinuity research design. Note that the rest of the empirical question will use the file 2miledata.dta. The housing variables in this file are 2 mile averages.**


## Part (a)
**Consider the HRS score as the running variable for an RD research design. What assumptions are needed on the HRS score? How do each of the following "facts" impact the appropriateness of these assumptions:**

### ------ a.i
**The EPA assertion that \the 28.5 cutooff was selected because it produced a manageable number of sites." **



### ------ a.ii
**None of the individuals involved in identifying the site, testing the level of pollution, or running the 1982 HRS test knew the cutoff threshold score. **



### ------ a.iii
**EPA documentation emphasizes that the HRS test is an imperfect scoring measure. **








\newpage
## Part (b)
**Create a histogram of the distribution (i.e. density) of the 1982 HRS scores by dividing the HRS score into non-overlapping bins. Include a vertical line at 28.5. Next, run local linear regressions on either side of 28.5 using the midpoints of the bins as the data. What do you conclude?**













\newpage
<!--=========================================================================-->
# Problem 3
<!--=========================================================================-->
**This question examines the 1st stage equation of an RD design using the 1982 HRS score.**


## Part (a)
**Use a 2SLS (IV) econometric setup that uses whether or not a census tract has a site scoring above/below 28.5 as the instrument. Write down the 1st stage equation. Run the 1st stage regression experimenting with the same set of covariates used in question (1). In addition, run a second specification in which you limit the sample to only those census tracts with sites between 16.5 and 40.5 and run the specification using all of the control variables (we will use this as the size of the bandwidth for the "regression discontinuity" regression). Interpret the results.**


## Part (b)
**Create a graph plotting the the 1982 HRS score against whether a site is listed on the NPL by year 2000 (NPL on the y-axis, HRS on the x -axis). Briefly explain and interpret this graph.**


## Part (c)
**Create a graph that plots the 1982 HRS score against 1980 property values (property values on the y-axis, HRS on the x -axis). What do you conclude from this graph?**






\newpage
<!--=========================================================================-->
# Problem 4
<!--=========================================================================-->
**Write down the 2nd stage equation (with housing values as the out-come) and the 2 standard assumptions for valid IV estimation. Run 2SLS to get the estimated coefficient on 2000 NPL status. Run the same two specifications as in the previous question. Briefly interpret the results.**








\newpage
<!--=========================================================================-->
# Problem 5
<!--=========================================================================-->
**Write a 1 paragraph conclusion summarizing your findings and interpreting the results. Be sure to comment on how the evidence from this problem set supports the primary research question.**








\newpage
# Appendix A: R Code

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```





















